---
title: "732A90 - Lab 6 - jorva845"
author: "Joris van Doorn - jorva845"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: pdf_document
---

```{r setup, include=FALSE}
RNGversion(min(as.character(getRversion()), "3.6.2"))
knitr::opts_chunk$set(echo = TRUE)
library(RMaCzek)
library(knitr)
library(tidyr)
library(tidyverse)
library(tinytex)
library(dplyr)
library(readxl)
library(stats)
library(coda)
library(gdata)
```

# Q1: Genetic Algorithm

*In this assignment, you will try to perform one-dimensional maximization with the help of a genetic algorithm.*

## 1.

*Define the function*

$$f(x) := \frac{x^2}{e^2} - 2 e^{-\frac{9 sin(x)}{x^2+x+1}}$$

```{r}
foo <- function(x){
  result <- (x^2 / exp(x))-2 * exp((-9*sin(x))/(x^2+x+1))
  return(result)
}
```

## 2.

*Define the function crossover(): for two scalars x and y it returns their \kid" as (x+y)=2.*

```{r}
crossover <- function(x, y){
  kid <- (x + y)/2
  return(kid)
}
```

## 3.

*Define the function mutate() that for a scalar x returns the result of the integer division x^2 mod 30. (Operation mod is denoted in R as %%).*

```{r}
mutate <- function(x){
  mod <- x^2 %% 30
  return(mod)
}
```

## 4.

*Write a function that depends on the parameters maxiter and mutprob and:*

### a.

*Plots function f in the range from 0 to 30. Do you see any maximum value?*


```{r}
genetic <- function(maxiter, mutprob){
  # a
  func_out <- c()
  for(i in 1:30){
    func_out[i] <- foo(i)
  }
  graph <- plot(1:30+maxiter, func_out, col = "black")
  
  # b
  init_pop <- seq(0,30,5)
  
  # c
  values <- c()
  for(i in 1:length(init_pop)){
    values[i] <- foo((i-1)*5)
  }
  
  max_val <- c()
  
  # d i
  for(i in 1:maxiter){
    parents <- sample(1:length(init_pop), 2, replace = F)
    
    # d ii
    victim <- order(values)[1]
    
    # d iii
    kid <- crossover(parents[1], parents[2])
    mutated_kid <- mutate(kid)
    
    # d iv
    init_pop[victim] <- mutated_kid
    
    for(j in 1:length(init_pop)){
      values[j] <- foo(init_pop[j])
    }
    
    max_val[i] <- max(init_pop)
    
    # d v
    graph
    points(30+i, max_val[i], col = "blue")
  }
}


```

### b.

*Defines an initial population for the genetic algorithm as X = (0, 5, 10, 15, ..., 30).*

### c.

*Computes vector Values that contains the function values for each population point.*

### d. 

*Performs maxiter iterations where at each iteration*

#### i. 

*Two indexes are randomly sampled from the current population, they are further used as parents (use sample()).*

#### ii.

*One index with the smallest objective function is selected from the current population, the point is referred to as victim (use order()).*

#### iii.

*Parents are used to produce a new kid by crossover. Mutate this kid with probability mutprob (use crossover(), mutate()).*

#### iv.

*The victim is replaced by the kid in the population and the vector Values is updated.*

#### v.

*The current maximal value of the objective function is saved.*

### e.

```{r}
genetic(10,0.1)
genetic(100, 0.1)
genetic(10, 0.5)
genetic(100, 0.5)
genetic(10, 0.9)
genetic(100, 0.9)
```

\newpage

# Q2: EM algorithm

\newpage

# Appendix

```{r, ref.label=knitr::all_labels(),echo=TRUE,eval=FALSE,results='show'}
```

